Supermartingales and Termination Analysis of Statistical PCF

### 2021-01-21 LICS21/related.tex
Add a simplified non-affine and continuous example.
Related work: added reference to more complex non-affine recursion example.

Kobayashi et al.~\cite{DBLP:conf/lics/KobayashiLG19} show that the termination probability of order-$n$ probabilistic recursion schemes ($n$-PHORS) can be obtained as the least fixpoint of suitable order-$(n-1)$ fixpoint equations, which are solvable by Kleene fixpoint iteration.  
By contrast, our approach works for programs with continuous distributions. 
Note that $n$-PHORS is strictly less expressive then order-$n$ call-by-name PPCF:
% (because the underlying recursion schemes are not Turing complete):
the PPCF terms in \cref{ex:raven complex,ex:higher-order recursion}
%\refExample{complexExample} 
are not definable as PHORS.

The main result by McIver et al.~\cite{DBLP:journals/pacmpl/McIverMKK18} is similar to our \cref{thm:antitone rankable implies termination}, but in an imperative language. While we require that the antitone ranking function decreases for every $\tY$-reduction step, they require that it decreases for every iteration of a certain while loop (the ranking function in that case is defined in the context of a particular loop), which is similar to $\tY$ but limited to tail recursion. 
The difference in the exact progress condition is not significant. 
\akr{Their progress condition, that there are antitone functions $p$ and $d$ such that the ranking function decreases from $x$ by at least $d(x)$ with probability at least $p(x)$, would not have worked directly because it does not cope as easily with the fact that the ranking function may change in value at the intermediate non-$\tY$ reduction steps (provided it does not increase in expectation). However, it is easy enough to convert a ranking function that satisfies either progress condition to the other.}
\lo{Can you replace the highlighted by a brief comment on their discussion of completeness?}

>>>>

Kobayashi et al.~\cite{DBLP:conf/lics/KobayashiLG19} show that the termination probability of order-$n$ probabilistic recursion schemes ($n$-PHORS) can be obtained as the least fixpoint of suitable order-$(n-1)$ fixpoint equations, which are solvable by Kleene fixpoint iteration.  
By contrast, our approach works for programs with continuous distributions. 
Note that $n$-PHORS is strictly less expressive then order-$n$ call-by-name PPCF:
% (because the underlying recursion schemes are not Turing complete):
the PPCF terms in \cref{ex:raven complex,ex:non-affine continuous,ex:higher-order recursion}
%\refExample{complexExample} 
are not definable as PHORS.

The main result by McIver et al.~\cite{DBLP:journals/pacmpl/McIverMKK18} is similar to our \cref{thm:antitone rankable implies termination}, but in an imperative language. While we require that the antitone ranking function decreases for every $\tY$-reduction step, they require that it decreases for every iteration of a certain while loop (the ranking function in that case is defined in the context of a particular loop), which is similar to $\tY$ but limited to tail recursion. 
The difference in the exact progress condition is insignificant. A ranking function satisfying either progress condition can easily be converted to satisfy the other.
\akr{Their progress condition, that there are antitone functions $p$ and $d$ such that the ranking function decreases from $x$ by at least $d(x)$ with probability at least $p(x)$, would not have worked directly because it does not cope as easily with the fact that the ranking function may change in value at the intermediate non-$\tY$ reduction steps (provided it does not increase in expectation). However, it is easy enough to convert a ranking function that satisfies either progress condition to the other.}
\lo{Can you replace the highlighted by a brief comment on their discussion of completeness?}
\akr{I would, but there's no space. Their completeness results seem pretty weak anyway.}
